---
title: "PShaji_Assignment7"
author: "Priya Shaji"
date: "10/12/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Let X1, X2, . . . , Xn be n mutually independent random variables, each of which is uniformly distributed on the integers from 1 to k. Let Y denote the minimum of the Xi’s. Find the distribution of Y.

## Answer 1

Let us compute the function of Y

Given that:

Y = Min(X1, X2, . . . , Xn) = Xi 

where  Xi is the minimum value.

We can find the distribution of Y by applying binomial distribution.

Therefore,

$n!/(n-1)!1! * p^(1) * q^(n-1)$

$=  n * p * q^(n-1)$

$=  n *  (1/k)* (n-1/k)^(n-1)$

$=  n * (1/k) * (n-1)^(n-1)/k^(n-1)$

$=  n * (n-1)^(n-1)/K^n$

Where,

$Xi = successful trial (minimum value)$

$X1 to Xi-1, Xi+1 to Xn = failed trials$

$n = number of trials$

$k = population size (total possible outcomes per trial)$

$p = 1/k probability of success$

$q = n-1/k probability of failure$

## Question 2

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).

a. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

b. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

c. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)

d. What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.


## Answer 2

<strong>Ans a)</strong>
Given: Expects one failure every ten yrs

Mean(machine_fails) = 10 yrs

Now, for geometric model distribution:

mean = 1/p

10 = 1/p

Probability of failure:

p = 1/10


1) Standard deviation:

```{r}
p = 0.10
SD = sqrt((1-p)/(p^2))
SD
```


Probability that the machine will fail after 8 years:

Discrete:

```{r}
p = 0.10
count = 0
for (i in 1:8)
{
   count = count + (1-p)^i*p
}
1 - count
```

<strong>Ans b)</strong>

Mean(machine_fails) = 10 yrs

Exponential probability distribution mean :

= beta = 1/lambda = 1/10

standard deviation, for exponential distribution:

= 1/lamba = 1/10

The cummlative distribution fucntion (CDF) for an exponetial distribution:

$f(x,lambda) = 1 - e^(-lambdax), where x >= 0 and 0 where x < 0 f(8,1/10)= 1 - e^(-1/10(8)), for x = 8$

Probability that the machine will fail after 8 years:

Continuous:

```{r}
count = (1 - exp(1)^(-0.10*8))
1 - count

```


<strong>Ans c)</strong>

Probability of success in 8 years:

p = 1/10

Number of trials = 8

Standard deviation:

```{r}
p = 0.10
n = 8
SD = sqrt(n*p*(1-p))
SD
```


k = number of successes in n trials = 0 PMF

$Formula for binomial distribution - P(k,n.p) = p(X=0) = nCrK * p^k * (1-p)^n-k$

Probability of 8 straight failures without success

```{r}

p = 0.10
n = 8
k =0
count = choose(n, k)*p^k*(1-p)^(n-k)
count
```


<strong>Ans d)</strong>


Poisson distribution:

lambda = rate/unit time = 1

machine breakdown per 10 years = expected value

Standard deviation is sqrt(lambda) = sqrt(1) = 1

Probability of k = 0 breakdowns in years 1 through 8

Probability of 8 straight failures without success

```{r}
p = 0.10
n = 8
k =0 
count = (1^k*exp(1)^(-1))/factorial(k)  
count

```






