---
title: "PShaji_Assignment9"
author: "Priya Shaji"
date: "10/21/2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### The price of one share of stock in the Pilsdorff Beer Company (see Exer- cise 8.2.12) is given by Yn on the nth day of the year. Finn observes that the differences Xn = Yn+1 − Yn appear to be independent random variables with a common distribution having mean μ = 0 and variance σ2 = 1/4. If Y1 = 100, estimate the probability that Y365 is (a) ≥ 100. (b) ≥ 110. (c) ≥ 120.

<strong>Answer</strong>

<strong>(a) ≥ 100 </strong>

```{r}
mean <- 0  
var <- 1/4
sd <- sqrt(var)
n <- 364  
x <- 0/sqrt(n)
pnorm(x, mean, sd, lower.tail = FALSE)
```


<strong>(b) ≥ 110 </strong>

```{r}
x <- 10/sqrt(n)
pnorm(x, mean, sd, lower.tail = FALSE)
```

<strong>(b) ≥ 120 </strong>

```{r}
x <- 20/sqrt(n)
pnorm(x, mean, sd, lower.tail = FALSE)

```


### Calculate the expected value and variance of the binomial distribution using the moment generating function.

<strong>Answer</strong>

Binomial moment generating function:

$M(t) =  (pe^t + q)^n$

Expected value => 1st derivative

$M`(t) = n(pe^t + q)^{n-1}* pe^t$

$E(x) = M`(0) = np$

Variance => 2nd derivative

$M``(t) = n[1-p+pe^t]^{n-1}(pe^t)n(n-1)(1-p+pe^t)^{n-2}(pe^t)$

$E(x^2) = M``(0) = n(n-1) p^2 + np$

$var(x) = E(x^2) - E(x)^2$

$var(x) = n(n-1)p^2+np-(np)^{2}$

$var(x) = (n^2p^2 - 1 np^2) +np - (np)^{2}$

$var(x) = (np^2) - np^2 +np - (np)^{2}$

$var(x) = np - np^2$

$var(x) = np(1-p)$

### Calculate the expected value and variance of the exponential distribution using the moment generating function.

<strong> Answer </strong>

Exponential moment generating function:

$M(t) = λ/λ-t$

Expected value => 1st derivative

$M`(t) = λ/(λ-t)^{2}$

$E(x) = M`(0) = λ/λ^2 = 1/λ$

Variance => 2nd derivative

$M``(t) = 2λ/(λ-t)^3$

$E(x^2) = M``(0) = 2λ/λ^3 = 2/λ^2$

$var(x) = E(x^2) - E(x)^2$

$var(x) = 2/λ^2 - 1/λ^2 = 1/λ^2$






















