---
html_document:
  df_print: paged
author: "Priya Shaji"
date: "10/20/2019"
output:
  html_document:
    df_print: paged
subtitle: Chapter 11,Page 424, Question 15
title: "HW 10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Consider the game of tennis when deuce is reached. If a player wins the next point, he has advantage. On the following point, he either wins the game or the game returns to deuce. Assume that for any point, player A has probability .6 of winning the point and player B has probability .4 of winning the point. (a) Set this up as a Markov chain with state 1: A wins; 2: B wins; 3: advantage A; 4: deuce; 5: advantage B. (b) Find the absorption probabilities. (c) At deuce, find the expected duration of the game and the probability that B will win.

<strong>Answer (a)

(a) Set this up as a Markov chain with state 1: A wins; 2: B wins; 3: advantage A; 4: deuce; 5: advantage B. </strong>

Step 1)  Compose transition matrix


```{r}
Markov_chain_state_names <- c("S1: Game_Win A", "S2: Game_Win B", "S3: Advan A", "S4: Deuce", "S5: Advan B")
Markov_chain<- matrix(c(1, 0, 0, 0, 0,
                  0, 1, 0, 0, 0,
                  .6, 0, 0, .4, 0,
                  0, 0, .6, 0, .4,
                  0, .4, 0, .6, 0),
                nrow = 5, byrow = TRUE)
colnames(Markov_chain) <- Markov_chain_state_names
rownames(Markov_chain) <- Markov_chain_state_names
Markov_chain
```



<strong> Answer (b)

(b) Find the absorption probabilities. </strong>

Step 2)  Change the transition matrix to canonical form

```{r}

Markov_chain <- Markov_chain[, c(3:5, 1:2)]
Markov_chain <- Markov_chain[c(3:5, 1:2),]
Markov_chain
```


Step 3) Subset matrix Q (transient to transient) 

Step 4) Subset matrix R (transient to absorbing)

Step 5) Compose an identity matrix I with same dimensions as Q

Step 6) Compute the fundamental matrix by solving the set of linear equations

Step 7) Compute absorption probabilities

```{r}
Q <- Markov_chain[1:3, 1:3]  
R <- Markov_chain[1:3, 4:5]  
I <- diag(3) 
N <- solve(I - Q)  
M <- N %*% R 
M
```


<strong>Answer (c)

(c) At deuce, find the expected duration of the game and the probability that B will win </strong>

Step 8) Compute expected steps to absorption

Step 9) Column vector of 1s

Step 10) Calculate expected steps to absorption

Step 11) Expected duration of the game in steps

```{r}

c<- c(rep(1, 3))   
Nc <- N %*% c  


Nc[2,1]
```

Step 12) Probability that B will win


```{r}

M[2,2]
```




































