---
title: "PShaji_Assignment10"
author: "Priya Shaji"
date: "11/2/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

### Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6. Find the probability that he wins 8 dollars before losing all of his money if (a) he bets 1 dollar each time (timid strategy). (b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy). (c) Which strategy gives Smith the better chance of getting out of jail?

<strong> Answer  (a) he bets 1 dollar each time (timid strategy) </strong>

probability of winning A dollars: 0.4

probability of losing A dollars: 0.6

```{r}
p <- 0.4
q <- 0.6
r <- q/p


for (i in 0:8){
ans <- round((1-r^i)/(1-r^8),4)
print (ans)
}
```


<strong> Answer (b) he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy) </strong>

```{r}
bold_1 <- matrix(c(1,0.6,0.6,0.6,0,0,0,0,0,0,0,0.4,0,0,0,0,0,0.4,0,0,0,0,0,0.4,1),nrow = 5, byrow=FALSE)
bold_1
```


```{r}
bold_2 <- matrix(c(0,1,0,0,0),nrow=1)
bold_2
```


```{r}
bold_3 <- bold_2 %*% bold_1
bold_3
```


```{r}
bold_4 <- bold_3 %*% bold_1
bold_5 <- bold_4 %*% bold_1
bold_5
```


Therefore, the probabity is 0.064


<strong> Answer (c) Which strategy gives Smith the better chance of getting out of jail?
 </strong>


The probability of bold strategy is better than timid strategy to get out of jail.







