---
title: "PShaji_Assignment5"
author: "Priya Shaji"
date: "9/27/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions.
Note that the point (B,C) is then chosen at random in the unit square.

Find the probability that

(a) B + C < 1/2

(b) BC < 1/2

(c) |B − C| < 1/2

(d) max{B,C} < 1/2

(e) min{B,C} < 1/2
 
 
## Answer 1)

1) Choose independently two numbers B and C at random from the interval [0, 1] with uniform density

```{r}
## Generating numbers at random from the interval [0, 1]

B <- runif(100000,min=0,max=1)

C<- runif(100000,min=0,max=1)

```

2) Prove that B and C are proper probability distributions.


```{r}

## Minimum value of B

min(B)

## Maximum value of B

max(B)

## Histogram of B

hist(B,probability = TRUE)
```



```{r}

## Minimum value of C

min(C)

## Maximum value of C

max(C)

## Histogram of C

hist(C,probability = TRUE)
```

By analyzing above graphs and simulaitons, we prove that B and C are proper probability distributions.

Frequency of B and Frequency of C are positive and also area under curve is 1.

(a) B + C < 1/2

Answer a)

```{r}
BC_a<- sum((B+C)< 0.5)/100000
BC_a
```



(b) BC < 1/2

Answer b)

```{r}
BC_b<- sum((B*C)<0.5)/100000
BC_b
```

(c) |B − C| < 1/2

Answer c)

```{r}
BC_c<- sum(abs(B-C)< 0.5)/100000
BC_c
```

(d) max{B,C} < 1/2

Answer d)

```{r}
BC_d<- sum(pmax(B,C)< 0.5)/100000
BC_d
```

(e) min{B,C} < 1/2

Answer e)

```{r}
BC_e<- sum(pmin(B,C)< 0.5)/100000
BC_e
```



