---
title: "PShaji_Assignment2"
author: "Priya Shaji"
date: "9/4/2019"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem set 1

(1) Show that $A^TA$ not equal to  $AA^T$ in general. (Proof and demonstration.)

Answer 1)

<strong> Demonstration </strong>

Let $A =$

\[
\begin{bmatrix}
a & b\\
c & d
\end{bmatrix}
\]

And, $A^T =$

\[
\begin{bmatrix}
e & f\\
g & h
\end{bmatrix}
\]


Therefore,

$A^TA =$


\[
\begin{bmatrix}
ea+fc & eb+fd \\
ga+hc & gb+hd
\end{bmatrix}
\]

Now, as we can see that the products of the $A$ and $A^t$ is not equal to each other.

<strong>Testing</strong>

```{r}
A <- matrix(c(10,11,12,13),ncol = 2)
t<- 2
ATA <- A^t %*% A
AAT <- A %*% A^t

identical(ATA, AAT)
```

Hence, using R function 'identical()' , it is proved that product of transpose of a matrix(AT) with the original matrix(A) i.e. $A^TA$,is not equal to the product of original matrix(A) with its transpose(AT), i.e. $AA^T$



(2) For a special type of square matrix A, we get $A^T A = AA^T$ . Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

Answer 2)

For a special type of square matrix A, we get $A^T A = AA^T$. 

The above statement is true provided the given matrices are symmetrical.

Let us see the demonstration in R:

1) Matrix $A$

```{r}
A <- matrix(c(1, 2, 1, 2, 1, 2, 1, 2, 1), nrow = 3, byrow = T)
A
```


2) Transpose of Matrix A: $A^T$


```{r}
t(A)
```


3) Test $A^TA = AA^T$

```{r}
ATA <- t(A) %*% A
AAT <- A %*% t(A)

AAT == ATA
```


Hence verified, the conditions required for a special type of square matrix A, we get $A^T A = AA^T$, is that the Matrix should be Symmetrical.


## Problem set 2

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

Answer)

1) Let's generate a sequence of random numbers using `set.seed`

2) set the size of matrix as n^2 , where n is number of elements in a row or column

```{r}
set.seed(605)
n <- 4
A <- matrix(sample.int(8,size = n^2, replace = TRUE), ncol = n)
A
```

3) Create a funcntion LU_matrix

4) for the L matrix , we will set the diagonal elements as 1.

5) Now we will perform row operations in `L` and `U` matrix and iterate through the matrix via `for()` loop

```{r}
LU_matrix <- function(A){
  
  U = A 
  L = diag(x = 1, ncol = ncol(A), nrow = nrow(A))
  
  for (i in 1:(nrow(U) -1)) {
    for (j in (i+1):nrow(U)){
      if (U[i, i] != 0){
        multiplier = U[j, i] / U[i, i]
        L[j, i] = multiplier
        U[j, ] = U[j, ] - multiplier * U[i, ]
      }
      
    }
    
  }
  return(list('L' = L, 'U' = U))
  
}

x <- LU_matrix(A)

## Print L matrix
x$L
```



```{r}
## Print U matrix

x$U
```

6) Perform matrix multiplication 

```{r}
x$L %*% x$U
```


As we can see that the above matrix is A matrix, therefore our L and U matrices produces verified results.


Now let us test our `LU_matrix()` function

```{r}
matrix_test <- matrix(c(1:9), ncol = 3)
test <- LU_matrix(matrix_test)
## Print matrix L
test$L
```


```{r}
## Print matrix U
test$U
```

Perform matrix multiplication to verify the result

```{r}
all.equal(test$L %*% test$U, matrix_test)
```


Therefore, product of matrix L and matrix U is equal to matrix `matrix_test`, which verifies the result.

